Compositional Reasoning about Connectors Using Coq and Z3

### 2017-06-19 FACS_Extension/paper_revised0508.tex
revised(lack of result)

\documentclass[3p,times]{elsarticle}
%\usepackage{ecrc}
\usepackage{graphicx}
% used for flow charts
\usepackage{listings}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{reotex}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

@@

\newcommand{\redt}[1]{\textcolor{red}{#1}}
\newcommand{\liyi}[1]{\textcolor{blue}{#1}}
%\newcommand{\xyz}[1]{\textcolor{blue}{#1}}

@@

\usepackage{url}
%\pagestyle{plain}
\begin{document}
\begin{frontmatter}
%\pagestyle{plain}
%\mainmatter

\title{Reasoning about Connectors in Coq}

@@

\author{Xiyue Zhang, Weijiang Hong, Yi Li and Meng Sun}

%\authorrunning{X.~Zhang, W.~Hong, Y.~Li and M.~Sun}

\address{Department of Informatics and LMAM, School of Mathematical Sciences,\\ Peking University}
%\email{\{zhangxiyue,wj.hong,liyi\_math,sunm\}@pku.edu.cn}}

%\maketitle
\begin{abstract}
Reo is a channel-based exogenous coordination model in which complex coordinators, called connectors, are compositionally
built out of simpler ones. In this paper, we present a new approach to model connectors in Coq which is a proof assistant based on higher-order logic and $\lambda$-calculus.
The model reflects the original structure of connectors simply and clearly. In our framework, basic connectors (channels) are interpreted as axioms and composition operations are specified as inference rules. Furthermore, connectors are interpreted as logical predicates which describe the relation
between inputs and outputs. With such definitions provided, connector properties, as well as equivalence and refinement relations between different connectors, can be naturally formalized as \emph{goals} in Coq and easily proved using pre-defined \emph{tactics}. Particularly, to fulfil the proof of refinement relation, Z3 is used as an assistant tool for solving logical formulas.
%\paragraph*{Keywords:}

\end{abstract}
\begin{keyword}
Coordination language, Reo, Coq, Reasoning, Z3
\end{keyword}
\end{frontmatter}
\section{Introduction}\label{sec:introduction}

@@

As an example, Reo \cite{Arb04}, which is a channel-based model for exogenous coordination, offers
a powerful language for implementation of coordinating component connectors. Connectors provide the protocols that control and organize
the communication, synchronization and cooperation among the components that they interconnect. Primitive connectors, called {\em channels}
in Reo, can be composed to build complex connectors. Reo has been successfully applied in different application domains, such as
service-oriented computing and bioinformatics \cite{CCA04,SA07}. In recent years,
verifying the correctness of connectors is becoming a critical challenge, especially due to the advent of Cloud computing
technologies. The rapid growth of size and complexity of the computing infrastructures has made it more difficult to model
and verify connector properties, and thus leads to less confidence on the correctness of connectors.

@@

In this paper, we aim to provide an approach to formally modeling and reasoning about connectors using Coq.
The basic idea of our
approach is to model the behavior of a connector by representing it as a logical predicate which describes the relation among
the timed data streams on the input and output nodes, and to reason about connectors' properties, as
well as the equivalence and refinement relations
between connectors, by using proof principles and tactics in Coq.
Compared with existing approaches for verifying connectors' properties \cite{BBK+10,KB09,KKV12}, using Coq is especially helpful when
we take infinite behavior into consideration. The coinductive proof principle makes it possible to prove connectors' properties easily while it
is difficult (sometimes impossible) for other approaches (like model checking) because of the huge (or maybe infinite) number of states. The main drawback of failing to check the complementary refinement relation in Coq is made up by Z3.
While Coq provides a platform to prove properties with a series of tactics and strategies to be considered, Z3 can be used as a component in the context of other tools that require solving logical formulas. The logical formulas involved in the model of Reo should reflect the constraints of connectors and be consistent with their behavior.

This is not a brand new idea, as we have already provided a solution for modeling Reo in Coq in\cite{LS15}, where connectors are represented in a constructive way, and
verification is essentially based on simulations. We do believe that the approach in this paper is reasonably different from
its predecessor where Coq seldom shows its real power. To be more specific, our new work has its certain
advantanges comparing with \cite{LS15} in the following
aspects:
\begin{itemize}
\item {\bf Modeling Method:} We use axioms to describe basic channels and their composition operations, which is more natural on a proof-assistant
platform than the simulation-based approach in \cite{LS15}.
\item {\bf Expression Power:} Any valid Coq expression can be used to depict properties, which is obviously more powerful than just using LTL formulas in
\cite{LS15}. Furthermore, support for continuous time behavior is also
possible in our approach in this paper.
\item {\bf Refinement and Equivalence Checking:} In our framework, equivalence and refinement relations can be proved among different connectors, while the previous one is not capable of either equivalence or refinement checking.
\end{itemize}

The paper is organized as follows: After this general introduction, we briefly summarize Reo, Coq and Z3
in Section \ref{sec:pre}. Section \ref{sec:basicdef} shows the notion of timed data streams and some
pre-defined auxiliary functions and predicates. Section \ref{sec:channelandoperator} presents the
formal modeling of basic channels and operators, as well as complex connectors. Section
\ref{sec:verification} shows how to reason about connector properties and equivalence (or
refinement) relations in our framework. Section \ref{sec:refinement} concentrates on the proof of refinement relation between two connectors.
In Section \ref{sec:conclusion}, we concludes with some further research directions. Full source codes can be found at \cite{reo2coqfile} for further reference.

@@

In this section, we provide a brief introduction to the coordination language Reo, proof assistant tools Coq and Z3.

\subsection{The Coordination Model Reo}
Reo is a channel-based exogenous coordination model wherein complex coordinators, called connectors,
are compositionally built out of simpler ones \cite{Arb04}.
Further details about Reo and its semantics can be found in \cite{Arb04,AR03,BSAR06}.
The simplest connectors are channels with well-defined behavior such as synchronous channels, FIFO channels, etc.
Each channel in Reo has exactly two directed ends, with their own identities.
There are two types of channel ends: source ends and sink ends. A source channel end accepts
data into the channel. A sink channel end dispenses data out of the channel.
\begin{figure}
  \centering
  \begin{tikzpicture}
  \draw (0.7, 0.6) node[draw=none] {Sync};
  \sync{(0,1)}{(1.5,1)}{}
  \draw (2.7, 0.6) node[draw=none] {Syncdrain};
  \syncdrain{(2,1)}{(3.5,1)}{node[below]{}}
  \draw (4.7, 0.6) node[draw=none] {FIFO1};
  \fifoe{(4,1)}{(5.5,1)}{node[below]{}}
  \draw (6.7, 0.6) node[draw=none] {Asyncdrain};
  \asyncdrain{(6,1)}{(7.5,1)}{node[below left]{}}
  \draw (8.7, 0.6) node[draw=none] {LossySync};
  \lossysync{(8,1)}{(9.5,1)}{node[below left]{}}
  \end{tikzpicture}\vspace{-6mm}
  \caption{Five types of basic channels.}\label{fig:basicchannel}
\end{figure}

The graphical notations of some basic channels are presented in Fig. \ref{fig:basicchannel}, and their behavior can be interpreted as follows:
\begin{itemize}
\item{\textbf{Sync:} a synchronous channel with one source end and one sink end. The pair of I/O operations on its two ends can succeed only simultaneously.}
\item{\textbf{SyncDrain:} a synchronous channel which has two source ends.
 The pair of input operations on its two ends can succeed only simultaneously.
 All data items written to this channel are lost.}
\item{\textbf{FIFO$n$:} an asynchronous channel with one source end and one sink end, and a bounded buffer with capacity $n$.
It can accept data items from its source end. The accepted data items are kept in the internal buffer, and dispensed to
the sink end in FIFO order.
Especially, the FIFO$1$ channel is an instace of FIFO$n$ where the buffer capacity is 1.}
\item{\textbf{AsyncDrain:}  an asynchronous channel} which has two source ends. The channel guarantees that the operations on its two ends never succeed simultaneously. All data items written to this channel are lost.
\item{\textbf{LossySync:} a synchronous channel with one source end
    and one sink end. The source end always accepts all data items. If
    there is no matching output operation on the sink end of the
    channel at the time that a data item is accepted, then the data
    item is lost; otherwise, the channel transfers the data item
    exactly the same as a Sync channel, and the output operation at the sink end succeeds.}
\end{itemize}

\begin{figure}
\centering
\begin{tikzpicture}
  \ionode{(io1)}{(0,5)}{}
  \ionode{(io2)}{(2,5)}{}
  \ionode{(io3)}{(4,4.7)}{}
  \ionode{(io4)}{(4,5.3)}{}
  \ionode{(io5)}{(8,4.7)}{}
  \ionode{(io6)}{(8,5.3)}{}
  \mixednode{(m1)}{(1,5)}{}
  \mixednode{(m2)}{(5,5)}{}
  \mixednode{(m3)}{(7,5)}{}
  \sync{(io1)}{(m1)}{}
  \sync{(m1)}{(io2)}{}
  \sync{(io3)}{(m2)}{}
  \sync{(io4)}{(m2)}{}
  \sync{(m3)}{(io5)}{}
  \sync{(m3)}{(io6)}{}
\end{tikzpicture}
\caption{Operations of channel composition.}\label{fig:channelcomposition}
\end{figure}

Complex connectors are constructed by composing simpler ones via the join and hiding operations. Channels are joined together in nodes. The set of channel ends coincident on a node is disjointly partitioned into the sets of source and sink channel ends that coincide on the node, respectively. Nodes are categorized into source, sink and mixed nodes, depending on whether all channel ends that coincide on a node are source ends, sink ends or a combination of the two. The hiding operation is used to hide the internal topology of a connector. The hidden nodes can no longer be accessed or observed from outside. There are three types of operations for channel composition: flow-through, merge and replicate. Fig. \ref{fig:channelcomposition} provides the graphical representation of these operations.

@@

Coq\cite{huet1997coq} is a widely-used proof assistant tool, where denotational formalizations (e.g.
theorem and hypothesis) and operational formalizations (e.g. functions and algorithms) are naturally
integrated. Moreover, it allows the interactive construction of formal proofs.
The formal language used in Coq is called \emph{Gallina}, which provides a convenient way to define
both programming statements and mathematical propositions, for example:
\begin{lstlisting}[language=coq]
  (* a variable definition *)
  Variables a b: nat.
  (* a simple non-recursive function *)
  Definition inc(a:nat) := a + 1.
  (* axioms don't have to be proved *)
  Axiom inc_ax: forall c:nat, inc(c) > c.
  (* theorems rely on proving *)
  Theorem inc_eq: forall c:nat, inc(c) = c + 1.
  Proof.
    (* interactive proving based on tactics *)
    auto.
  Qed.
\end{lstlisting}

@@

Z3\cite{MouraB08} is an efficient SMT Solver and a high performance theorem prover freely available from Microsoft Research. It can be used in various software verification and analysis applications. Acting as a SMT (Satisfiability Modulo Theories) solver, Z3 expands to deciding the satisfiability (or dually the validity) of first order formulas with respect to combinations of theories such as: arithmetic, bit-vectors, arrays, and uninterpreted functions. Given the data and time constraints of connectors, it allows us to decide the satisfiability against properties or refinement relations. Z3 provides bindings for several programming languages. In this paper, we use the Z3 API in Python to construct models and carry out experiments. The following example will provide an intuitive understanding of Z3 solver.
\begin{lstlisting}[frame=single]
x = Int('x')
y = Int('y')

@@

#Restoring state
s.pop()
#Solving restored set of constraints
print s.check()
\end{lstlisting}
This example contains two constraints at first. From these two constraints, we can see that Z3 supports Boolean operators, such as \emph{And}, \emph{Or}, \emph{Not}, \emph{Implies (implication)}, apart from the operators like $<$, $>$ for comparison.
The command \emph{Solver()} creates a general purpose solver. Constraints can be added using the method \emph{add}. After being added to the solver, the constraints can be said having been asserted in the solver . The method \emph{check()} solves the asserted constraints. Finally, the result is satisfiable (unsatisfiable) if a solution was (not) found, respectively. A solver may fail to solve a system of constraints and \emph{unknown} is returned. Each solver maintains a stack of assertions. The command \emph{push} creates a new scope by saving the current stack size. The command \emph{pop} removes any assertion performed between it and the matching push.
\section{Basic Definitions}\label{sec:basicdef}

@@

%\subsection{Timed data streams}
The behavior of a connector can be formalized by means of data-flows at its sink and source nodes which are essentially infinite sequences. With the help of the stream library in Coq, such infinite data-flows can be defined as \emph{timed data streams}:
\begin{lstlisting}[language=coq]
    Definition Time := R.
    Definition Data := nat.
    (*Inductive Data : Set :=
        |Natdata : nat-> Data
        |Empty : Data.*)
    Definition TD := Time * Data.
    Variable Input : Stream TD.
    Variable Output : Stream TD.
\end{lstlisting}
%
In our framework, time is represented by real numbers. Benefit from the completeness of real number system, we can express and carry out the effective operation of a quantity at any precision request.
The continuity of the set of real numbers is sufficiently enough for our modeling approach. Also the
continuous time model is more appropriate since it is very expressive and closer to the nature of time in the real world. Thus, the time sequence consists of increasing and diverging time moments. For simplicity, here we take the natural numbers as the definition of data, which can be easily expanded according to different application domains. The Cartesian product of time and data defines a TD object.
We use the stream module in Coq to produce streams of TD objects.

%\subsection{Auxiliary functions and predicates}
Some auxiliary functions and predicates are defined to facilitate the representation of axioms for basic channels in Reo. This part can be extended for further use in different problems.

The terms ``PrL" and ``PrR" take a pair of values (a, b) that has Cartesian product type A $\times$ B as the argument and return the first or second value of the pair, respectively.
%\begin{verbatim}
%    Definition PrL {A B: Type}(pair: A * B) := (...).
%    Definition PrR {A B: Type}(pair: A * B) := (...).
%\end{verbatim}

The following functions provide some judgment of time, which can make
the description of axioms and theorems for connectors more concise and
clear. ``Teq" means that time of two streams are equal and ``Tneq" has the
opposite meaning. ``Tle"  (``Tgt") represents that time of the first stream is
strictly less (greater) than the second stream. The judgement about
equality of data is analogous to the judgement of time. The complete
definition of these functions can be found at \cite{reo2coqfile}.

@@

A LossySync channel behaves the same as a Sync channel, except that a write operation on its source always succeeds immediately. If a compatible read or take operation is already pending on the sink of a LossySync channel, the written data item is transferred to the pending operation and both succeed. Otherwise, the write operation succeeds and the data item is lost. The LossySync channel can be defined as follows:
\begin{lstlisting}[language=coq]
   Parameter LossySync: Stream TD -> Stream TD -> Prop.
   Axiom LossySync_coind:
     forall Input Output: Stream TD,
     LossySync Input Output ->
       (
        (hd Output = hd Input  /\ LossySync (tl Input)(tl Output))
        \/
        LossySync(tl Input) Output
       ).
\end{lstlisting}

AsyncDrain is analogous to SyncDrain except that it guarantees that the pairs of write operations on the two channel ends never succeed simultaneously. Similarly it only has requirements on the time of the two streams on its opposite ends, but it requires that the times of the two streams are always different. The AsyncDrain channel can be defined as follows:
\begin{lstlisting}[language=coq]
    Parameter AsyncDrain: Stream TD -> Stream TD -> Prop.
    Axiom AsyncDrain_coind:
      forall Input1 Input2: Stream TD,
      AsyncDrain Input1 Input2 ->
        (~ PrL(hd Input1)  =  PrL (hd Input2) )
        /\
        ( (
          (PrL(hd Input1) < PrL (hd Input2)) /\
          AsyncDrain (tl Input1) Input2
          ) /\
          (
          (PrL(hd Input1) > PrL (hd Input2))  /\
          AsyncDrain Input1 (tl Input2)
          )
        ).
\end{lstlisting}

@@

\subsection{Derivation of Connector Properties}
The proof process of a property is as follows: the user states the proposition that needs to be proved, called a \emph{goal},
then he/she applies commands called \emph{tactics} to decompose this goal into simpler subgoals or solve it directly. This decomposition
process ends when all subgoals are completely solved. In the following, we use some examples to illustrate our approach instead of
giving all the complex technical details.
\begin{example}
We first consider the connector given in Fig. \ref{fig:compsyncfifo}, which consists of two channels \emph{AB} and \emph{BC} with types Sync and FIFO1, respectively.

We use $a$ and $b$ to denote the time streams when the corresponding data streams flow into and out of the Sync channel \emph{AB}, and
\emph{c} to denote the time stream for the data stream that flows out of the FIFO1 channel \emph{BC}. Here we can see that a flow-through
operation has acted on the mixed node \emph{B}. The time when the stream flows into the FIFO1 channel \emph{BC} is equal to the time when the
stream flows out of the Sync channel \emph{AB}. The following theorem states the property $a < c$ for this connector. The connector is based on the
axioms Sync and FIFO1, which can be used as hypotheses for the proof of the theorem.
\begin{theorem}\label{the:tleac}
$\forall A,B,C.\:Sync(A,B)\land FIFO1(B,C) \rightarrow Tle(A,C)$.
\end{theorem}

@@

\begin{figure}
\vspace{0cm}
\centering
\begin{tikzpicture}
\ionode{(io1)}{(-5,1.2)}{node[left]{A}}
\ionode{(io2)}{(-5,0)}{node[left]{B}}
\ionode{(io3)}{(-2,1.2)}{node[right]{C1}}
\ionode{(io4)}{(-2,0)}{node[right]{C2}}
\syncdrain{(io1)}{(io2)}{}
\fifoe{(io1)}{(io3)}{}
\sync{(io2)}{(io4)}{}
\draw (-3.5, -0.6) node[draw=none] {(a) Before merge};
%\draw (-1.6, -0.6) node[draw=none] {Transformed version of alternator.};

>>>>

%\documentclass[3p,times]{elsarticle}
%\usepackage{ecrc}
\documentclass[preprint,3p]{elsarticle}
\usepackage{graphicx}
% used for flow charts
\usepackage{listings}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{reotex}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

@@

\newcommand{\redt}[1]{\textcolor{red}{#1}}
%\newcommand{\liyi}[1]{\textcolor{blue}{#1}}
%\newcommand{\xy}[1]{\textcolor{blue}{#1}}

@@

\usepackage{url}

\journal{Science of Computer Programming}

%\pagestyle{plain}
\begin{document}
\begin{frontmatter}
%\pagestyle{plain}
%\mainmatter

\title{Compositional Reasoning about Connectors Using Coq and Z3}

@@

\author{Xiyue Zhang}\ead{zhangxiyue@pku.edu.cn}
\author{Weijiang Hong}\ead{wj.hong@pku.edu.cn}
\author{Yi Li}\ead{liyi$\_$math@pku.edu.cn}
\author{Meng Sun}\ead{sunmeng@math.pku.edu.cn}
\address{LMAM \& DI, School of Mathematical Science, Peking University, Beijing, China}
%\email{\{zhangxiyue,wj.hong,liyi\_math,sunm\}@pku.edu.cn}}

%\maketitle
\begin{abstract}
Reo is a channel-based exogenous coordination language in which complex coordinators, called connectors, are compositionally built out of simpler ones. In this paper, we present an approach to model and reason about connectors using Coq and Z3. %Coq is a proof assistant based on higher-order logic and $\lambda$-calculus.
The model reflects the original structure of connectors simply and clearly. In our framework, basic connectors (channels) are interpreted as axioms and composition operations are specified as inference rules. Furthermore, connectors are interpreted as logical predicates which describe the relations
between inputs and outputs. With such definitions provided, connector properties, as well as equivalence and refinement relations between different connectors, can be naturally formalized as \emph{goals} in Coq and easily proved using pre-defined \emph{tactics}. Particularly, to fulfill the proof of refinement relation, Z3 is used as an assistant tool for solving logical formulas.
%\paragraph*{Keywords:}

\end{abstract}
\begin{keyword}
Coordination language, Reo, Coq, Z3, Reasoning
\end{keyword}
\end{frontmatter}
\section{Introduction}\label{sec:introduction}

@@

As an example, Reo \cite{Arb04}, which is a channel-based exogenous coordination language, offers
a powerful framework for implementation of coordinating component connectors. Connectors provide the protocols that control and organize
the communication, synchronization and cooperation among the components that they interconnect. Primitive connectors, called {\em channels}
in Reo, can be composed to build complex connectors. Reo has been successfully applied in different application domains, such as
service-oriented computing and bioinformatics \cite{CCA04, SA07}. In recent years,
verifying the correctness of connectors is becoming a critical challenge, especially due to the advent of Cloud computing
technologies. The rapid growth of size and complexity of the computing infrastructures has made it more difficult to model
and verify connector properties, and thus leads to less confidence on the correctness of connectors.

@@

In a previous paper \cite{ZHL+17} we showed how to formally model and reason about connectors' properties in Coq. The basic idea of our approach is to model the behavior of a connector by representing it as a logical predicate which describes the relation among the timed data streams on the input and output nodes, and to reason about connectors' properties, as well as the equivalence and refinement relations between connectors, by using proof principles and tactics in Coq. Compared with existing approaches for verifying connectors' properties \cite{BBK+10,KB09,KKV12}, reasoning about connectors using theorem provers like Coq is especially helpful when we take infinite behavior into consideration. The coinductive proof principle makes it possible to prove connectors' properties easily while it is difficult (sometimes impossible) for other approaches (like model checking) because of the huge (or maybe infinite) number of states.

In the current paper, we aim to provide a more complete account of the approach in \cite{ZHL+17} to formally model and reason about connectors. The main drawback of failing to check the complementary refinement relation in Coq is made up by using the constraint solver Z3. While Coq provides a platform to prove properties with a series of tactics and strategies to be considered, Z3 can be used as a component in the context %of other tools
that requires solving logical formulas. The logical formulas involved in the model of Reo should reflect the constraints of connectors and be consistent with their behavior.

The approach is not a brand new idea, as we have already provided a solution for modeling Reo in Coq in \cite{LS15}, where connectors are represented in a constructive way, and verification is essentially based on simulations. We do believe that the approach in this paper is reasonably different from its predecessor \cite{LS15} where Coq seldom shows its real power. To be more specific, our new work has its certain
advantages comparing with \cite{LS15} in the following aspects:
\begin{itemize}
\item {\bf Modeling Method:} We use axioms to describe basic channels and their composition operations, which is more natural on a proof-assistant platform than the simulation-based approach in \cite{LS15}.
\item {\bf Expression Power:} Any valid Coq expression can be used to depict properties, which is obviously more powerful than just using LTL formulas in \cite{LS15}. Furthermore, support for continuous time behavior is also possible in our approach in this paper.
\item {\bf Refinement and Equivalence Checking:} In our framework, equivalence and refinement relations can be proved among different connectors, while the approach in \cite{LS15} is not capable of either equivalence or refinement checking. Furthermore, the constraint solver Z3 is adopted to prove non-refinement relations between connectors, while in \cite{ZHL+17} we can only derive proofs for refinement relations.
\end{itemize}

The paper is organized as follows: After this general introduction, we briefly summarize Reo, Coq and Z3
in Section \ref{sec:pre}. Section \ref{sec:basicdef} shows the notion of timed data streams and some
pre-defined auxiliary functions and predicates. Section \ref{sec:channelandoperator} presents the
formal modeling of basic channels and operators, as well as complex connectors. Section
\ref{sec:verification} shows how to reason about connector properties and equivalence (or
refinement) relations in our framework. Section \ref{sec:refinement} concentrates on the proof of refinement relation between two connectors.
In Section \ref{sec:conclusion}, we conclude with some further research directions. Full source codes can be found at \cite{reo2coqfile} for further reference.

@@

In this section, we provide a brief introduction to the coordination language Reo, and the proof assistant tools Coq and Z3.

\subsection{The Coordination Model Reo}
Reo is a channel-based exogenous coordination language wherein complex coordinators, called connectors,
are compositionally built out of simpler ones \cite{Arb04}.
Further details about Reo and its semantics can be found in \cite{Arb04, AR03, BSAR06}.
The simplest connectors are channels with well-defined behavior such as synchronous channels, FIFO channels, etc.
Each channel in Reo has exactly two directed ends, with their own identities.
There are two types of channel ends: \emph{source} ends and \emph{sink} ends. A source channel end accepts
data into the channel. A sink channel end dispenses data out of the channel.
\begin{figure}
  \centering
  \begin{tikzpicture}
  \draw (0.7, 0.6) node[draw=none] {Sync};
  \sync{(0,1)}{(1.5,1)}{}
  \draw (2.7, 0.6) node[draw=none] {Syncdrain};
  \syncdrain{(2,1)}{(3.5,1)}{node[below]{}}
  \draw (4.7, 0.6) node[draw=none] {FIFO1};
  \fifoe{(4,1)}{(5.5,1)}{node[below]{}}
  \draw (6.7, 0.6) node[draw=none] {LossySync};
  \lossysync{(6,1)}{(7.5,1)}{node[below left]{}}
  \end{tikzpicture}\vspace{-6mm}
  \caption{Four types of basic channels.}\label{fig:basicchannel}
\end{figure}

The graphical notations of some basic channels are presented in Fig.~\ref{fig:basicchannel}, and their behavior can be interpreted as follows:
\begin{itemize}
\item{\textbf{Sync:} a synchronous channel with one source end and one sink end. The pair of I/O operations on its two ends can succeed only simultaneously.}
\item{\textbf{SyncDrain:} a synchronous channel which has two source ends.
 The pair of input operations on its two ends can succeed only simultaneously.
 All data items written to this channel are lost.}
\item{\textbf{FIFO$n$:} an asynchronous channel with one source end and one sink end, and a bounded buffer with capacity $n$.
It can accept data items from its source end. The accepted data items are kept in the internal buffer, and dispensed to
the sink end in FIFO order.
Especially, the FIFO$1$ channel is an instace of FIFO$n$ where the buffer capacity is 1.}
%\item{\textbf{AsyncDrain:}  an asynchronous channel} which has two source ends. The channel guarantees that the operations on its two ends never succeed simultaneously. All data items written to this channel are lost.
\item{\textbf{LossySync:} a synchronous channel with one source end
    and one sink end. The source end always accepts all data items. If
    there is no matching output operation on the sink end of the
    channel at the time that a data item is accepted, then the data
    item is lost; otherwise, the channel transfers the data item
    exactly the same as a Sync channel, and the output operation at the sink end succeeds.}
\end{itemize}

\begin{figure}[ht]
\centering
\begin{tikzpicture}
  \ionode{(io1)}{(0,5)}{}
  \ionode{(io2)}{(2,5)}{}
  \ionode{(io3)}{(4,4.7)}{}
  \ionode{(io4)}{(4,5.3)}{}
  \ionode{(io5)}{(8,4.7)}{}
  \ionode{(io6)}{(8,5.3)}{}
  \mixednode{(m1)}{(1,5)}{}
  \mixednode{(m2)}{(5,5)}{}
  \mixednode{(m3)}{(7,5)}{}
  \sync{(io1)}{(m1)}{}
  \sync{(m1)}{(io2)}{}
  \sync{(io3)}{(m2)}{}
  \sync{(io4)}{(m2)}{}
  \sync{(m3)}{(io5)}{}
  \sync{(m3)}{(io6)}{}
\end{tikzpicture}
\caption{Operations of channel composition.}\label{fig:channelcomposition}
\end{figure}

Complex connectors are constructed by composing simpler ones via the join and hiding operations. Channels are joined together in nodes. The set of channel ends coincident on a node is disjointly partitioned into the sets of source and sink channel ends that coincide on the node, respectively. Nodes are categorized into source, sink and mixed nodes, depending on whether all channel ends that coincide on a node are source ends, sink ends or a combination of the two. The hiding operation is used to hide the internal topology of a connector. The hidden nodes can no longer be accessed or observed from outside. There are three types of operations for channel composition: \emph{flow-through}, \emph{merge} and \emph{replicate}. Fig.~\ref{fig:channelcomposition} provides the graphical representation of these operations.

@@

Coq\cite{huet1997coq} is a widely-used proof assistant tool, where denotational formalizations (e.g.
theorem and hypothesis) and operational formalizations (e.g. functions and algorithms) are naturally
integrated. Moreover, it allows the interactive construction of formal proofs.
The formal language used in Coq, \emph{Gallina}, provides a convenient way to define
both programming statements and mathematical propositions, for example:
\begin{lstlisting}[language=coq]
  (* a variable definition *)
  Variables a b: nat.
  (* a simple non-recursive function *)
  Definition inc(a:nat) := a + 1.
  (* axioms don't have to be proved *)
  Axiom inc_ax: forall c:nat, inc(c) > c.
  (* theorems rely on proving *)
  Theorem inc_eq: forall c:nat, inc(c) = c + 1.
  Proof.
    (* interactive proving based on tactics *)
    auto.
  Qed.
\end{lstlisting}

@@

Z3 \cite{MouraB08} is an efficient SMT (Satisfiability Modulo Theories) Solver and a high performance theorem prover freely available from Microsoft. It can be used in various software verification and analysis applications. Acting as a SMT solver, Z3 expands to deciding the satisfiability (or dually the validity) of first order formulas with respect to combinations of theories such as: arithmetic, bit-vectors, arrays, and uninterpreted functions. Given the data and time constraints of connectors, it allows us to decide the satisfiability against properties or refinement relations. Z3 provides bindings for several programming languages. In this paper, we use the Z3 API in Python to construct the models and carry out experiments. The following example provides an intuitive understanding of the Z3 solver.
\begin{lstlisting}[frame=single]
x = Int('x')
y = Int('y')

@@

#Restoring state
s.pop()
#Solving restored set of constraints
print s.check()
\end{lstlisting}

This example contains two constraints at first. From these two constraints, we can see that Z3 supports Boolean operators, such as \emph{And}, \emph{Or}, \emph{Not} and \emph{Implies (implication)}, apart from the operators like $<$, $>$ for comparison.
The command \emph{Solver()} creates a general purpose solver. Constraints can be added using the method \emph{add}. Once they are added to the solver, we can say that the constraints have been asserted in the solver. The method \emph{check()} solves the asserted constraints. Finally, the result is satisfiable (unsatisfiable) if a solution is (not) found, respectively. A solver may fail to solve a system of constraints and \emph{unknown} is returned. Each solver maintains a stack of assertions. The command \emph{push} creates a new scope by saving the current stack size. The command \emph{pop} removes any assertion performed between it and the matching \emph{push}.
\section{Basic Definitions}\label{sec:basicdef}

@@

%\subsection{Timed data streams}
The behavior of a connector can be formalized by means of data-flows at its sink and source nodes which are essentially infinite sequences. With the help of the stream library in Coq, such infinite data-flows can be defined as \emph{timed data streams}:
\begin{lstlisting}[language=coq]
    Definition Time := R.
    Definition Data := nat.
    (*Inductive Data : Set :=
        |Natdata : nat-> Data
        |Empty : Data.*)
    Definition TD := Time * Data.
    Variable Input : Stream TD.
    Variable Output : Stream TD.
\end{lstlisting}
%
In our framework, time is represented by real numbers. %Benefit from the completeness of real number system, we can express and carry out the effective operation of a quantity at any precision request.
The continuity of the set of real numbers is sufficiently enough for our modeling approach. Also the
continuous time model is more appropriate than the discrete model since it is very expressive and closer to the nature of time in the real world. Thus, the time sequence consists of increasing and diverging time moments. For simplicity, here we take the natural numbers as the definition of data, which can be easily expanded according to different application domains. The Cartesian product of time and data defines a TD object.
We use the stream module in Coq to produce streams of TD objects.

%\subsection{Auxiliary functions and predicates}
Some auxiliary functions and predicates are defined to facilitate the representation of axioms for basic channels in Reo. This part can be extended for further use in different problems.
The terms ``PrL" and ``PrR" take a pair of values (a, b) that has Cartesian product type A $\times$ B as the argument and return the first or second value of the pair, respectively.
%\begin{verbatim}
%    Definition PrL {A B: Type}(pair: A * B) := (...).
%    Definition PrR {A B: Type}(pair: A * B) := (...).
%\end{verbatim}
The following functions provide some judgment of time, which can make
the description of axioms and theorems for connectors more concise and
clear: ``Teq" means that time of two streams are equal and ``Tneq" has the
opposite meaning. ``Tle"  (``Tgt") represents that time of the first stream is
strictly less (greater) than the second stream. The judgement about
equality of data is analogous to the judgement of time. The complete
definition of these functions can be found at \cite{reo2coq2Z3}.

@@

A LossySync channel behaves the same as a Sync channel, except that a write operation on its source end always succeeds immediately. If a compatible read or take operation is already pending on its sink end, the written data item is transferred to the pending operation and both succeed. Otherwise, the write operation succeeds and the data item is lost. The LossySync channel can be defined as follows:
\begin{lstlisting}[language=coq]
   Parameter LossySync: Stream TD -> Stream TD -> Prop.
   Axiom LossySync_coind:
     forall Input Output: Stream TD,
     LossySync Input Output ->
       (
        (hd Output = hd Input  /\ LossySync (tl Input)(tl Output)) \/
        LossySync(tl Input) Output
       ).
\end{lstlisting}

AsyncDrain is analogous to SyncDrain except that it guarantees that the pairs of write operations on the two channel ends never succeed simultaneously. Similarly it only has requirements on the time of the two streams on its opposite ends, but it requires that the times of the two streams are always different. The AsyncDrain channel can be defined as follows:
\begin{lstlisting}[language=coq]
    Parameter AsyncDrain: Stream TD -> Stream TD -> Prop.
    Axiom AsyncDrain_coind:
      forall Input1 Input2: Stream TD,
      AsyncDrain Input1 Input2 ->
        (~ PrL(hd Input1)  =  PrL (hd Input2) ) /\
        (
          ( (PrL(hd Input1) < PrL (hd Input2)) /\ AsyncDrain (tl Input1) Input2) /\
          ( (PrL(hd Input1) > PrL (hd Input2))  /\ AsyncDrain Input1 (tl Input2) )
        ).
\end{lstlisting}

%Note that in the above definitions we choose Axiom to define LossySync and AsyncDrain
%because it is easier to use the coinductive expression to specify their behavior.

@@

\subsection{Derivation of Connector Properties}
The proof process of a property is as follows: the user states the proposition that needs to be proved, called a \emph{goal},
then he/she applies commands called \emph{tactics} to decompose this goal into simpler subgoals or solve it directly. This decomposition
process ends when all subgoals are completely solved. In the following, we use some examples to illustrate our approach instead of
giving all the complex technical details.
\begin{example}
We first consider the connector given in Fig. \ref{fig:compsyncfifo}, which consists of two channels \emph{AB} and \emph{BC} with types Sync and FIFO1, respectively.
We use $a$ and $b$ to denote the time streams when the corresponding data streams flow into and out of the Sync channel \emph{AB}, and
$c$ to denote the time stream for the data stream that flows out of the FIFO1 channel \emph{BC}. Here we can see that a flow-through
operation has acted on the mixed node \emph{B}. The time when the stream flows into the FIFO1 channel \emph{BC} is equal to the time when the
stream flows out of the Sync channel \emph{AB}. The following theorem states the property $a < c$ for this connector. The connector is based on the
axioms Sync and FIFO1, which can be used as hypotheses for the proof of the following theorem.
\end{example}

\begin{theorem}\label{the:tleac}
$\forall A,B,C.\:Sync(A,B)\land FIFO1(B,C) \rightarrow Tle(A,C)$.
\end{theorem}

@@

In Coq, the theorem can be proved as follows. Note that the formalism is slightly different from the previous one. By the \emph{section} environment, Coq is able to encapsulate hypothesizes as assumptions of the theorem. So the two definitions are exactly equivalent.
